### 2022-03-23 paper/localpsf.tex
minor changes

The direct approach to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ is to apply $\Aop$ to a point source centered at $x$ to get $\impulseresponse_x$, as per \eqref{eq:impulse_response_delta_action}. Then one can post-process $\impulseresponse_x$ to determine $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$. But this direct approach is computationally infeasible because it requires one operator action of $\Aop$ per point $x$, and our method will need to know $V(x)$, $\mu(x)$, and $\Sigma(x)$ at all points $x$. Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$, \emph{for all points $x\in\Omega$ simultaneously}, by applying $\Aop^T$ to one constant function, $\gdim$ linear functions, and $\gdim(\gdim+1)/2$ quadratic functions. Specifically, let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^\gdim\right)$, and let $C$, $\{L^i\}_{i=1}^\gdim$, and ${\{Q^{ij}\}_{i=1}^\gdim}_{j=1}^\gdim$ be the following constant, linear, and quadratic functions:
\begin{equation*}
C(x) := 1, \qquad
L^i(x) := x^i, \qquad
Q^{ij}(x) = x^i x^j.
\end{equation*}
The following results (derived in Appendix \ref{app:proofs}) follow straightforwardly from the definitions:
\begin{subequations}
	\label{eq:vol_mean_var}
	\begin{align}
	\spatialvol =& \left(\Aop^T C\right)^* \\
	\spatialmean^i =& \left(\Aop^T L^i\right)^* / \spatialvol \\
	\spatialcov^{ij} =& \left(\Aop^T Q^{ij}\right)^* / \spatialvol - \spatialmean^i\cdot \spatialmean^j
	\end{align}
\end{subequations}
for $i=1,\dots, \gdim$, $j=1,\dots,\gdim$. Here $f/g$ denotes pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ denotes pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$. The resulting method form computing $V$, $\mu$, and $\Sigma$ is shown in Algorithm \ref{alg:varhpi_mean_cov}.

\begin{algorithm2e}
	\SetAlgoNoLine
	\SetKwInOut{Input}{Input}
	\SetKwInOut{Output}{Output}
	{	
		\Input{Operator $\Aop$}
		\Output{$\spatialvol$, $\spatialmean$, and $\spatialcov$}

		\tcp{Compute scaling factor $\spatialvol$}

		Form constant function $C(x)=1$

		Compute $\spatialvol = \left(\Aop^T C\right)^*$

		\tcp{Compute mean $\spatialmean$}
		\For{$i=1,2,\dots,\gdim$}{
			Form linear function $L^i(x) = x^i$

			Compute $\spatialmean^i = \left(\Aop^T L^i\right)^* / \spatialvol$
		}
		\tcp{Compute covariance $\spatialcov$}
		\For{$i=1,2,\dots,\gdim$}{
			\For{$j=1,\dots,i$}{
				Form quadratic function $Q^{ij}(x) = x^i x^j$

				Compute $\spatialcov^{ij} = \left(\Aop^T Q^{ij}\right)^* / \spatialvol - \spatialmean^i\cdot \spatialmean^j$

				Set $\spatialcov^{ji} = \spatialcov^{ij}$

@@

	}
	\caption{Compute scaling factor $\spatialvol$, mean $\spatialmean$, and covariance $\spatialcov$}
	\label{alg:varhpi_mean_cov}
\end{algorithm2e}

@@

		Compute $\spatialvol$, $\spatialmean$, and $\spatialcov$ using Algorithm \ref{alg:varhpi_mean_cov}.

@@

We start with a finite set of candidate points $P$ and build $\pointbatch_k$ incrementally with points selected from $P$. For simplicity of explanation, here $\pointbatch_k$ and $P$ are mutable sets that we add points to and remove points from. First we initialize $\pointbatch_k$ as an empty set. Then we select the candidate point $p \in P$ that is the farthest away from all points in previous sample point batches $S_1 \cup S_2 \cup \dots \cup S_{k-1}$. Candidate points for the first batch $S_1$ are chosen randomly from $P$.
Once $p$ is selected, we remove $p$ from $P$. Then we perform the following two checks:
\begin{enumerate}
	\item We check whether $p$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_p \cap E_q = \{\}$ for all $q \in \pointbatch_k$.
	\item We make sure that $V(p)$ is not too small, by checking whether $V(p) > \tau V_\text{max}$. Here $V_\text{max}$ is the largest value of $V(q)$ over all points $q$ in the initial set of candidate points, and $\tau$ is a small threshold parameter (we use $\tau=10^{-5}$).
\end{enumerate}
If $p$ passes both these checks (i.e., if $p$ is sufficiently far from other points in the batch, and $V(p)$ is not too small) then we add $p$ to $\pointbatch_k$. Otherwise we discard $p$. This process repeats until there are no more points in $P$.  This is detailed in Algorithm \ref{alg:point_choice}. We check whether $E_p \cap E_q = \{\}$ using the ellipsoid intersection test described in Appendix \ref{sec:fast_ellipsoid_intersection_test}.

>>>>

The direct approach to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ is to apply $\Aop$ to a point source centered at $x$ to get $\impulseresponse_x$, as per \eqref{eq:impulse_response_delta_action}. Then one can post-process $\impulseresponse_x$ to determine $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$. But this direct approach is computationally infeasible because it requires one operator action of $\Aop$ per point $x$, and our method will need to know $V(x)$, $\mu(x)$, and $\Sigma(x)$ at all points $x$. Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$, \emph{for all points $x\in\Omega$ simultaneously}, by applying $\Aop^T$ to one constant function, $\gdim$ linear functions, and $\gdim(\gdim+1)/2$ quadratic functions. Specifically, let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^\gdim\right)$, and let $C$, $\{L^i\}_{i=1}^\gdim$, and ${\{Q^{ij}\}_{i=1}^\gdim}_{j=1}^\gdim$ be the following constant, linear, and quadratic functions:
\begin{equation*}
C(x) := 1, \qquad
L^i(x) := x^i, \qquad
Q^{ij}(x) = x^i x^j.
\end{equation*}
The following results (derived in Appendix \ref{app:proofs}) follow straightforwardly from the definitions:
\begin{subequations}
	\label{eq:vol_mean_var}
	\begin{align}
	\spatialvol =& \left(\Aop^T C\right)^* \\
	\spatialmean^i =& \left(\Aop^T L^i\right)^* / \spatialvol \\
	\spatialcov^{ij} =& \left(\Aop^T Q^{ij}\right)^* / \spatialvol - \spatialmean^i\cdot \spatialmean^j
	\end{align}
\end{subequations}
for $i=1,\dots, \gdim$, $j=1,\dots,\gdim$. Here $f/g$ denotes pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ denotes pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$. The resulting method for computing discretized versions of $V$, $\mu$, and $\Sigma$ is shown in Algorithm \ref{alg:varhpi_mean_cov}.

\begin{algorithm2e}
	\SetAlgoNoLine
	\SetKwInOut{Input}{Input}
	\SetKwInOut{Output}{Output}
	{	
		\Input{Operator $\bm{A}^T$}
		\Output{$\bm{\spatialvol}\in \mathbb{R}^\fedim$, $\bm{\spatialmean}\in \mathbb{R}^{\gdim \times \fedim}$, and $\bm{\spatialcov} \in \mathbb{R}^{\gdim \times \gdim \times \fedim}$}

		$\bm{\spatialvol} \gets \bm{M}^{-1}\bm{A}^T \bm{C}$

			$\bm{\spatialmean}^i \gets \left(\bm{M}^{-1}\bm{A}^T \bm{L}^i\right) / \bm{\spatialvol}$
		}
		\For{$i=1,2,\dots,\gdim$}{
			\For{$j=1,\dots,i$}{
				Form quadratic function $\bm{Q}^{ij} = x^i x^j$

				$\bm{\spatialcov}^{ij} \gets \left(\bm{M}^{-1}\bm{A}^T \bm{Q}^{ij}\right) / \bm{\spatialvol} - \bm{\spatialmean}^i\cdot \bm{\spatialmean}^j$

				$\bm{\spatialcov}^{ji} \gets \bm{\spatialcov}^{ij}$

@@

	}
	\caption{Compute discretized scaling factor $\mathbf{V}$, mean $\boldsymbol{\mu}$, and covariance $\boldsymbol{\Sigma}$}
	\label{alg:varhpi_mean_cov}
\end{algorithm2e}

%\begin{algorithm2e}
%	\SetAlgoNoLine
%	\SetKwInOut{Input}{Input}
%	\SetKwInOut{Output}{Output}
%	{	
%		\Input{Operator $\Aop$}
%		\Output{$\spatialvol$, $\spatialmean$, and $\spatialcov$}
%		
%		\tcp{Compute scaling factor $\spatialvol$}
%		
%		Form constant function $C(x)=1$
%		
%		Compute $\spatialvol = \left(\Aop^T C\right)^*$
%		
%		\tcp{Compute mean $\spatialmean$}
%		\For{$i=1,2,\dots,\gdim$}{
%			Form linear function $L^i(x) = x^i$
%			
%			Compute $\spatialmean^i = \left(\Aop^T L^i\right)^* / \spatialvol$
%		}
%		\tcp{Compute covariance $\spatialcov$}
%		\For{$i=1,2,\dots,\gdim$}{
%			\For{$j=1,\dots,i$}{
%				Form quadratic function $Q^{ij}(x) = x^i x^j$
%				
%				Compute $\spatialcov^{ij} = \left(\Aop^T Q^{ij}\right)^* / \spatialvol - \spatialmean^i\cdot \spatialmean^j$
%				
%				Set $\spatialcov^{ji} = \spatialcov^{ij}$
%				
%			}
%		}
%		
%	}
%	\caption{Compute scaling factor $\spatialvol$, mean $\spatialmean$, and covariance $\spatialcov$}
%	\label{alg:varhpi_mean_cov}
%\end{algorithm2e}

@@

		Compute $\bm{\spatialvol}$, $\bm{\spatialmean}$, and $\bm{\spatialcov}$ using Algorithm \ref{alg:varhpi_mean_cov}.

@@

We start with a finite set of candidate points $P$ and build $\pointbatch_k$ incrementally with points selected from $P$. For simplicity of explanation, here $\pointbatch_k$ and $P$ are mutable sets that we add points to and remove points from. First we initialize $\pointbatch_k$ as an empty set. Then we select the candidate point $p \in P$ that is the farthest away from all points in previous sample point batches $S_1 \cup S_2 \cup \dots \cup S_{k-1}$. Candidate points for the first batch $S_1$ are chosen randomly from $P$.
Once $p$ is selected, we remove $p$ from $P$. Then we perform the following two checks:
\begin{enumerate}
	\item We check whether $p$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_p \cap E_q = \{\}$ for all $q \in \pointbatch_k$.
	\item We make sure that $V(p)$ is not too small, by checking whether $V(p) > \epsilon V_\text{max}$. Here $V_\text{max}$ is the largest value of $V(q)$ over all points $q$ in the initial set of candidate points, and $\tau$ is a small threshold parameter (we use $\epsilon=10^{-5}$).
\end{enumerate}
If $p$ passes both these checks (i.e., if $p$ is sufficiently far from other points in the batch, and $V(p)$ is not too small) then we add $p$ to $\pointbatch_k$. Otherwise we discard $p$. This process repeats until there are no more points in $P$.  This is detailed in Algorithm \ref{alg:point_choice}. We check whether $E_p \cap E_q = \{\}$ using the ellipsoid intersection test described in Appendix \ref{sec:fast_ellipsoid_intersection_test}.
